---
title: "data_translation_project"
author: "Tamlyn Tamura"
date: "2/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
Loading libraries
Ctrl + Alt + I for new code chunks

```{r libraries, include=FALSE}
library(tidyverse)
library(jtools)
library(car)
library(vtable)
library(purrr)
```

Read in data

```{r data, include = FALSE}
id_name_link <- read_csv("Lab3_Rawdata/id_name_link.csv")

Most_Recent_Cohorts_Scorecard_Elements_ <- read_csv("Lab3_Rawdata/Most+Recent+Cohorts+(Scorecard+Elements).csv")

trends <- list.files(path = 'Lab3_Rawdata', pattern = 'trends_up_to', full.names = TRUE) %>%
  map_dfr(read_csv)
```

This is where I take unstructured data and try to make it a bit easier to read. I linked 'unitid' and 'opeid' to create a unique key, in case it became more organized this way down the line. 

```{r data wranging, include = FALSE}
id_name_link <- id_name_link %>% distinct(schname, .keep_all = TRUE)
id_name_link <- id_name_link %>% unite(key, c('unitid', 'opeid'), sep = '_')


Most_Recent_Cohorts_Scorecard_Elements_ <- Most_Recent_Cohorts_Scorecard_Elements_ %>%
  unite(key, c('UNITID', 'OPEID'), sep = '_')


college_id <- left_join(id_name_link, Most_Recent_Cohorts_Scorecard_Elements_)
college_id <- separate(college_id, schname, into = 'college', sep = '-')
college_id <- college_id[!duplicated(college_id$'college'),] # Separated college from cities then removed colleges with same names
college_id <- filter(college_id, PREDDEG == '3') # Filter to only show colleges that predominantly grant bachelor's degrees


college_id_condensed <- select(college_id, c('key', 'college', 'INSTNM', 'STABBR', 'PREDDEG', 'md_earn_wne_p10-REPORTED-EARNINGS', 'LOCALE'))


college_trends <- full_join(college_id_condensed, trends) %>%
  group_by(college, .add = FALSE) %>%
  subset(monthorweek > "2015-08-23 - 2015-08-29") # Filter out data that's from before September 2015, before College Scorecard was released

```

Create new data sets separated by LOCALE type--typically urbanized/city areas have a higher market rate than rural and suburban/town areas. 

I was hesitant on whether or not I should include Rural Fringe, Rural Distant, and Rural Remote (41, 42, 43), since there aren't many colleges that would be in these areas, but this actually accounted for 6448 observations, so I decided to keep it in.


```{r separate data sets, include = FALSE}

# Urban data frame
urban_df <- filter(college_trends, LOCALE == 11 & 12 & 13)
urban_df$`md_earn_wne_p10-REPORTED-EARNINGS` <- as.numeric(as.character(urban_df$`md_earn_wne_p10-REPORTED-EARNINGS`))
summary(urban_df)


# Suburban data frame
suburban_df <- filter(college_trends, LOCALE == 21 & 22 & 23)
suburban_df$`md_earn_wne_p10-REPORTED-EARNINGS` <- as.numeric(as.character(suburban_df$`md_earn_wne_p10-REPORTED-EARNINGS`))
summary(suburban_df)

# Town data frame
town_df <- filter(college_trends, LOCALE == 31 & 32 & 33)
town_df$`md_earn_wne_p10-REPORTED-EARNINGS` <- as.numeric(as.character(town_df$`md_earn_wne_p10-REPORTED-EARNINGS`))
summary(town_df)

# Rural data frame
rural_df <- filter(college_trends, LOCALE == 41 & 42 & 43)
rural_df$`md_earn_wne_p10-REPORTED-EARNINGS`<- as.numeric(as.character(rural_df$`md_earn_wne_p10-REPORTED-EARNINGS`))
summary(rural_df)

```

Urban min: 19800
Urban median: 46600
Urban mean: 47318
Urban max: 121500

Suburban min: 17600
Suburban median: 43900
Suburban mean: 45017
Suburban max: 166200

Town min: 28900
Town median: 39900
Town mean: 41275
Town max: 68500

Rural min: 21300
Rural median: 37800
Rural mean: 37375
Rural max: 59000

Despite a couple of outliers, it was expected that the more urbanized your surrounding is, the higher the median salary of someone 10 years out of college (1. Urban, 2. Suburban, 3. Town, 4. Rural)

```{r}

```

