---
title: "data_translation_project"
author: "Tamlyn Tamura"
date: "2/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/Desktop/School/MSBA/Stats 1 for Business Analytcs (R)/omsba_5112/omsba_5300_data_translation_project/Lab3_rawdata')
getwd()
```

## R Markdown
Loading libraries
Ctrl + Alt + I for new code chunks

```{r, include=FALSE}
knitr::opts_knit$set(root.dir = '../Lab3_rawdata')
getwd()
library(tidyverse)
library(jtools)
library(car)
library(vtable)
library(purrr)
library(ggplot2)
```

Read in data

```{r, include = FALSE}
knitr::opts_knit$set(root.dir = './Code/DTP_assessment.Rmd')
getwd()

id_name_link <- read_csv("id_name_link.csv")

Most_Recent_Cohorts_Scorecard_Elements_ <- read_csv("Most+Recent+Cohorts+(Scorecard+Elements).csv")

trends <- list.files(path = 'Lab3_Rawdata', pattern = 'trends_up_to', full.names = TRUE) %>%
  map_dfr(read_csv)
```

This is where I take unstructured data and try to make it a bit easier to read. I linked 'unitid' and 'opeid' to create a unique key, in case it became more organized this way down the line. 

```{r, include = FALSE}
knitr::opts_knit$set(root.dir = '../Code/DTP_assessment.Rmd')
getwd()

id_name_link <- id_name_link %>% distinct(schname, .keep_all = TRUE)
id_name_link <- id_name_link %>% unite(key, c('unitid', 'opeid'), sep = '_')


Most_Recent_Cohorts_Scorecard_Elements_ <- Most_Recent_Cohorts_Scorecard_Elements_ %>%
  unite(key, c('UNITID', 'OPEID'), sep = '_')


college_id <- left_join(id_name_link, Most_Recent_Cohorts_Scorecard_Elements_)
college_id <- separate(college_id, schname, into = 'college', sep = '-')
college_id <- college_id[!duplicated(college_id$'college'),] # Separated college from cities then removed colleges with same names
college_id <- filter(college_id, PREDDEG == '3') # Filter to only show colleges that predominantly grant bachelor's degrees


college_id_condensed <- select(college_id, c('key', 'college', 'INSTNM', 'STABBR', 'PREDDEG', 'md_earn_wne_p10-REPORTED-EARNINGS', 'LOCALE'))


college_trends <- full_join(college_id_condensed, trends) %>%
  group_by(college, .add = FALSE) %>%
  subset(monthorweek > "2015-08-23 - 2015-08-29") # Filter out data that's from before September 2015, before College Scorecard was released

```

Create new data sets separated by LOCALE type--typically urbanized/city areas have a higher market rate than rural and suburban/town areas. 

I was hesitant on whether or not I should include Rural Fringe, Rural Distant, and Rural Remote (41, 42, 43), since there aren't many colleges that would be in these areas, but this actually accounts for 6448 observations, so I have decided to leave those in.


```{r, include = FALSE}
knitr::opts_knit$set(root.dir = '../Code/DTP_assessment.Rmd')
getwd()

college_trends$`md_earn_wne_p10-REPORTED-EARNINGS` <- as.numeric(as.character(college_trends$`md_earn_wne_p10-REPORTED-EARNINGS`))
summary(college_trends) # This should change characters to numeric

# Urban data frame
urban_df <- filter(college_trends, LOCALE == 11 & 12 & 13, na.rm = TRUE)
summary(urban_df)


# Suburban data frame
suburban_df <- filter(college_trends, LOCALE == 21 & 22 & 23, na.rm = TRUE)
summary(suburban_df)


# Town data frame
town_df <- filter(college_trends, LOCALE == 31 & 32 & 33, na.rm = TRUE)
summary(town_df)


# Rural data frame
rural_df <- filter(college_trends, LOCALE == 41 & 42 & 43, na.rm = TRUE)
summary(rural_df)


```

Locale median: 42300
Locale mean: 43904

Urban min: 19800, 
Urban median: 46600
Urban mean: 47318
Urban max: 121500

Suburban min: 17600
Suburban median: 43900
Suburban mean: 45017
Suburban max: 166200

Town min: 28900
Town median: 39900
Town mean: 41275
Town max: 68500

Rural min: 21300
Rural median: 37800
Rural mean: 37375
Rural max: 59000

Urban environments mean there are more people, more infrastructure, more businesses, and a lot more money to go around, thus raising the market rate in the local area. Despite a couple of outliers, it was expected to notice that the more urbanized a college is located, the higher the mean and medians of the salary among those who graduated 10 years before (1. Urban, 2. Suburban, 3. Town, 4. Rural). 

Comparing these results to the median of total colleges across the board, I get a median 42300 USD and a mean of 43904 USD. These numbers are not too far off from the suburban mean and median, which is too be expected since there are significantly more colleges that predominantly grant bachelor's degrees (more observations) in the urban and suburban areas, as opposed to in town and rural areas.

Since market rates in more populated areas are higher than less populated areas, I will be defining high-earning colleges as colleges located in urban and suburban areas, while low-earning colleges will be those located in town and rural areas.


```{r, include = FALSE}

# Realized md_earn_wne_pt10-REPORTED-EARNINGS was causing me problems while running my regressions due to the -
# Went back and renamed the variable for my 5 data frames (urban_df, suburban_df, town_df, rural_df, locale_df)
locale_df <- rename(college_trends, median_salary = 'md_earn_wne_p10-REPORTED-EARNINGS')
urban_df <- rename(urban_df, urban_median_salary = median_salary)
suburban_df <- rename(suburban_df, suburban_median_salary = median_salary)
town_df <- rename(town_df, town_median_salary = median_salary)
rural_df <- rename(rural_df, rural_median_salary = median_salary)


locale_reg <- lm(index ~ median_salary, data = locale_df)
summary(locale_reg)

urban_reg <- lm(index ~ urban_median_salary, data = urban_df)
summary(urban_reg)

suburban_reg <- lm(index ~ suburban_median_salary, data = suburban_df)
summary(suburban_reg)

town_reg <- lm(index ~ town_median_salary, data = town_df)
summary(town_reg)

rural_reg <- lm(index ~ rural_median_salary, data = rural_df)
summary(rural_reg)
```

When I ran the regressions separately, I noticed that all the intercepts produced statistically significant numbers, except for in the town data frame. The rural data frame produced a median salary coefficient that was significant at the .01 level. The urban and suburban areas produced median salary coefficients significant at the .001 significance level.

When I ran the regression for all index on all locales, I got the median salary to be significant at the .001 level. What was interesting is that my Adjusted R^2 is ~.002, which means that this model doesn't tell us very much about the variation in the data, which is fair since it's very, very rare to make the conclusion that one variable that will drive the search popularity, or index.

Since the salary coefficients for urban and suburban locales were statistically significant at the 99.9% level, then it's safe to say that among colleges that predominantly grant bachelor's degrees, higher-earning colleges result in more student interest. Given the negative coefficient for median salaries for colleges in rural areas, we can say we are 99% confident that among the colleges that predominantly grant bachelor's degrees, lower-earning colleges and student interest have a negative relationship; meaning that lower-earning colleges are linked to less student interest. Additionally, we don't have statistical significance to prove whether or not colleges in towns are linked to higher or lower student interest.

Below I used ggplot to determine if my data is normally distributed. As you can see, it looks like there are some problems in my data given the shape. One possibility could be some of the outliers, since my data ranged from as low as 17,600 USD to 166,200 USD. Another possibility is that this is not a linear relationship, and I should have used a log-linear, linear-log or log-log model to display this relationship.

```{r}
g <- ggplot(locale_reg, aes(x=median_salary)) + geom_density()
print(g)

```


